Adaption of the Chumbley Score to matching of bullet striation marks

### 2018-01-03 doc_v2/doc_v2.Rmd
.

```{r missings, fig.width=8, fig.height=5, fig.cap="Number of failed tests by the window optimization size, wo, and ground truth. Test results from different sources have a much higher chance to fail, raising the question, whether failed tests should be treated as rejections of the null hypothesis of same source."}
ns <- all %>% group_by(wv, wo, match) %>% summarize(
  n_ss = mean(same_shift_n),
  n_ds = mean(diff_shift_n),
  zero_ss = sum(same_shift_n==0),
  zero_ds = sum(diff_shift_n==0),
  miss = sum(is.na(p_value)),
  missperc = miss/85491*100
)

#ns1<- ns[which(ns$wv == c(30,50), arr.ind = T),]
ns %>% filter(wv %in% c(30, 50)) %>%
#ns %>% filter(wv == c(30, 50)) %>%
  ggplot(aes(x = wo, y = missperc, shape=factor(wv), colour = factor(match))) + 
  geom_smooth(method="lm", size=.5, se=FALSE) +
  geom_point() + 
  theme_bw() +
  ylab("Percent of failed tests.") +
  xlab("Size of optimization window, wo") +
  scale_colour_brewer("Match", palette="Set1") +
  scale_shape_discrete("Size of \nvalidation\nwindow, wv")
```
Figure \ref{fig:missings} gives an overview of the number of failed tests, i.e. tests in which a particular parameter setting did not return a valid result. This happens, when the shift to align two signatures is so large, that the remaining overlap is too small to accommodate  windows for validation. The problem is therefore exacerbated by a larger validation window. Figure \ref{fig:missings} also shows that the number of failed tests is approximately linear in the size of the optimization window.

@@

total1<- errors_sig_c %>% ggplot(aes(x = coarse, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + geom_line()  + geom_point(data= errors2 %>% filter(wv==30), aes(shape="profiles")) + geom_line(data=errors2 %>% filter(wv==30), aes(shape="profiles")) + theme_bw() +
  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + xlab("Coarseness") +ylab("Total error")

total2<- wo120 %>% filter(wv>10) %>% ggplot(aes(x = wv, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + geom_line() +geom_point(data=errors2 %>% filter(wv>10) %>% filter(coarse==0.25), aes(shape="profiles")) + geom_line(data=errors2 %>% filter(wv>10)%>% filter(coarse==0.25), aes(shape="profiles")) + theme_bw() +
  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + ylab("Total error")

@@

#ns1<- ns[which(ns$wv == c(30,50), arr.ind = T),]
miss_1<- nsp %>% filter(wv == 30) %>%
#ns %>% filter(wv == c(30, 50)) %>%
  ggplot(aes(x = coarse, y = missperc, shape= "profile", colour = factor(match))) + 
  #geom_smooth(method="auto", size=.5, se=FALSE) +
  geom_point(size = 2) + 
  theme_bw() +
  ylab("Percent of failed tests.") +
  xlab("Coarseness") +
  scale_colour_brewer("Match", palette="Set1") + geom_point(data=ns_sig_c  %>% filter(wv %in% c(30,50)) , aes(shape="signature"), size = 2)# + scale_colour_brewer("Match", palette="Set2") #+geom_smooth(method="auto", size=.5, se=FALSE) 
  #scale_shape_discrete("Size of \nvalidation\nwindow, #wv")

miss_2<- nsp %>% filter(coarse == 0.25) %>%
#ns %>% filter(wv == c(30, 50)) %>%
  ggplot(aes(x = wv, y = missperc, shape= "profile", colour = factor(match))) + 
  #geom_smooth(method="auto", size=.5, se=FALSE) +
  geom_point(size = 2) + 
  theme_bw() +
  ylab("Percent of failed tests.") +
  xlab("wv") +
  scale_colour_brewer("Match", palette="Set1") + geom_point(data=ns %>% filter(wo == 120)  %>% filter(coarse == 1) , aes(shape="signature"), size = 2) #+ scale_colour_brewer("Match", palette="Set2")

>>>>

```{r missings, fig.width=8, fig.height=5, fig.cap="Number of failed tests by the window optimization size, wo, and ground truth."}
ns <- all %>% group_by(wv, wo, match) %>% summarize(
  n_ss = mean(same_shift_n),
  n_ds = mean(diff_shift_n),
  zero_ss = sum(same_shift_n==0),
  zero_ds = sum(diff_shift_n==0),
  miss = sum(is.na(p_value)),
  missperc = miss/85491*100
)

#ns1<- ns[which(ns$wv == c(30,50), arr.ind = T),]
ns %>% filter(wv %in% c(30, 50)) %>%
#ns %>% filter(wv == c(30, 50)) %>%
  ggplot(aes(x = wo, y = missperc, shape=factor(wv), colour = factor(match))) + 
  geom_smooth(method="lm", size=.5, se=FALSE) +
  geom_point() + 
  theme_bw() +
  ylab("Percent of failed tests.") +
  xlab("Size of optimization window, wo") +
  scale_colour_brewer("Match", palette="Set1") +
  scale_shape_discrete("Size of \nvalidation\nwindow, wv")
```
Figure \ref{fig:missings} gives an overview of the number of failed tests, i.e. tests in which a particular parameter setting did not return a valid result. This happens, when the shift to align two signatures is so large, that the remaining overlap is too small to accommodate  windows for validation. The problem is therefore exacerbated by a larger validation window. Figure \ref{fig:missings} also shows that the number of failed tests is approximately linear in the size of the optimization window. Test results from different sources have a much higher chance to fail, raising the question, whether failed tests should be treated as rejections of the null hypothesis of same source. For known non-matches there is a higher possibility that in the optimization pair of windows where cross-correlations are maximum are too far apart, and same shifts of this order hit the end of the signature.

Another reason for failed tests can be incorrect identification of maximum correlation windows in the optimization step as seen in \ref{fig:prof_missings}(c) because of the level of smoothing, as too much smoothing would subdue intricate features that might otherwise help in the correlation calculations and correct identification of maximum correlation windows irrespective of the size. This would again cause a simiar effect as explained for \ref{fig:missings} with validation windows, irrespective of size, during the shifts end up at the ends of the markings resulting in an invalid calculation and failed comparison attempt.

In \ref{fig:prof_missings}(c) and (d), we compare profiles and signatures on the basis of number of failed tests. The profiles chosen for \ref{fig:prof_missings}(d) have a constant coarseness of 0.25 and window of optimization as 120. The signatures in this case are not smoothed using the chumbley algorithm step of LOWESS smoothing. Instead signatures are used as calculated by \citet{aoas}. The smoothing in these signatures were determined and fixed on the basis of their performance in the random forest based algorithm proposed by \citet{aoas}. The comparison of profiles and signatures with variation of validation window size therefore is made on even footing. The trends are similar to \ref{fig:missings} in the sense that for known non-matches the number of failed tests are more for both signatures and profiles and increasing linearly with the validation window size. The problem is however, worse for profiles which has higher number of failed tests than signatures for all validation windows.

@@

total1<- errors_sig_c %>% ggplot(aes(x = coarse, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + geom_line()  + geom_point(data= errors2 %>% filter(wv==30), aes(shape="profiles")) + geom_line(data=errors2 %>% filter(wv==30), aes(shape="profiles")) + theme_bw() +
  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + xlab("Coarseness") +ylab("Total error") + ggtitle(label = "(a)")

total2<- wo120 %>% filter(wv>10) %>% ggplot(aes(x = wv, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + geom_line() +geom_point(data=errors2 %>% filter(wv>10) %>% filter(coarse==0.25), aes(shape="profiles")) + geom_line(data=errors2 %>% filter(wv>10)%>% filter(coarse==0.25), aes(shape="profiles")) + theme_bw() +
  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + ylab("Total error") + ggtitle(label = "(b)")

@@

#ns1<- ns[which(ns$wv == c(30,50), arr.ind = T),]
miss_1<- nsp %>% filter(wv == 30) %>%
#ns %>% filter(wv == c(30, 50)) %>%
  ggplot(aes(x = coarse, y = missperc, shape= "profile", colour = factor(match))) + 
  #geom_smooth(method="auto", size=.5, se=FALSE) +
  geom_point(size = 2) + 
  theme_bw() +
  ylab("Percent of failed tests.") +
  xlab("Coarseness") +
  scale_colour_brewer("Match", palette="Set1") + geom_point(data=ns_sig_c  %>% filter(wv %in% c(30,50)) , aes(shape="signature"), size = 2) + ggtitle(label = "(c)")
# + scale_colour_brewer("Match", palette="Set2") #+geom_smooth(method="auto", size=.5, se=FALSE) 
  #scale_shape_discrete("Size of \nvalidation\nwindow, #wv")

miss_2<- nsp %>% filter(coarse == 0.25) %>%
#ns %>% filter(wv == c(30, 50)) %>%
  ggplot(aes(x = wv, y = missperc, shape= "profile", colour = factor(match))) + 
  #geom_smooth(method="auto", size=.5, se=FALSE) +
  geom_point(size = 2) + 
  theme_bw() +
  ylab("Percent of failed tests.") +
  xlab("wv") +
  scale_colour_brewer("Match", palette="Set1") + geom_point(data=ns %>% filter(wo == 120)  %>% filter(coarse == 1) , aes(shape="signature"), size = 2) + ggtitle(label = "(d)")#+ scale_colour_brewer("Match", palette="Set2")
